---
title: "Reddy2016_EGFRTMT_sectimecourse"
author: "Alexis Chang"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
EGF100nM <- read_csv(file = "raw_data/MainFig3to6_SuppFig9to15/EGF100nM.csv") %>% 
  clean_names() %>% 
  # separate(sites, into = c("mod_site1", "mod_site2"), sep = " ", remove = FALSE) %>% 
  # pivot_longer(cols = c("mod_site1", "mod_site2"), names_to = "mod_site_num", values_to = "mod_site" ) %>% 
  # filter(!is.na(mod_site)) %>% #remove rows if only one psite per isoform
    mutate(mod_res = str_extract_all(sites, "[STY]"),
         mod_loc = str_remove_all(sites, "[pSTY]"),
         direction_1min = case_when(
           x60s - x0s > 0 ~ "up",
           x60s - x0s < 0~ "down",
           TRUE ~ "unchanged"))%>%
  # mutate(mod_res = str_sub(sites, start = 2L, end = 2L),
  #        mod_loc = as.numeric(str_sub(sites, start = 3L)),
  #        direction_1min = case_when(
  #          x60s - x0s > 0 ~ "up",
  #          x60s - x0s < 0~ "down",
  #          TRUE ~ "unchanged"))%>% 
  filter(grepl("Y", mod_res) == TRUE) #isoform must contain at least one pY site


#intermediate df for joining zscore + stdev---------------------------
EGF100nM_longer_zscore <- EGF100nM %>% 
  pivot_longer(cols = c(x0s:x80s), names_to = "time_point", values_to = "zscore") %>% 
  select(-c(x0s_stdev:x80s_stdev))

EGF100nM_longer_SD <-  EGF100nM %>% 
  pivot_longer(cols = c(x0s_stdev:x80s_stdev), names_to = "time_point", values_to = "stdev") %>% 
  select(-c(x0s:x80s)) %>% 
  mutate(time_point = str_sub(time_point, end = -7L ))


#master df for plotting --------------------------------
EGF100nM_longer <- EGF100nM_longer_zscore %>% 
  left_join(y = EGF100nM_longer_SD)  %>% 
  mutate(time_point_numeric = as.numeric(str_remove_all(time_point, "[xs]")),
         gene_ref = paste(protein_short," ", str_remove_all(sites, "[p]"), sep = "" ))





```
###__i. function to iterate through my data
```{r}
lineplot_gene_psite_function <- function(vector_gene_refs, dataframe, output_dir = "output/MainFig3to6_SuppFig9to15/Reddy2016Comparison/", condition = condition, sample_id = sample_id, strip_color = "gray80", strip_text_color = "black"){

  #create output directory if it doesn't exist already
  if (!dir.exists(output_dir)){
    dir.create(output_dir)}
  
  for (gene_var in vector_gene_refs) {
    if(gene_var %in% dataframe$gene_ref) {
      
      #create boxplot for any sites in each gene
      p <- ggplot() + 
        geom_line(data = dataframe %>% filter(gene_ref == gene_var),
                      mapping = aes(x = time_point_numeric, y = zscore)) +
        geom_errorbar(data = dataframe %>% filter(gene_ref == gene_var),
                      mapping = aes(x = time_point_numeric, y = zscore, ymax = zscore+stdev, ymin = zscore-stdev)) +
        # geom_boxplot(data = dataframe %>% filter(gene_ref == gene_var),
        #        mapping = aes(x = condition, y = median_norm_intensity), show.legend = FALSE, alpha = 0.3, outlier.color = NA, width = 0.5) +
        # geom_point(data = dataframe %>% filter(gene_ref == gene_var),
        #            mapping = aes(x = condition, y = median_norm_intensity),
        #            fill = "grey",
        #            mod_residue = position_jitter( width = 0.15, height = 0), shape = 21, alpha = 0.5, size = 1.5, show.legend = FALSE) +
        # geom_text(data = dataframe %>%
        # filter(gene_ref == gene_var) %>%
        #   mutate(max_intensity_condition = max(median_norm_intensity)) %>% 
        # distinct(condition, gene_ref, n_obs,max_intensity_condition),
        # aes(x = condition, y = max_intensity_condition + 1, label = n_obs),size = 4, show.legend = FALSE, fontface = "bold") +
        alexis_theme() +
        
        facet_wrap(facets = vars(gene_ref), nrow = 4) +
        scale_y_continuous(name = expression("scaled intensity"), expand = c(0.05,0.2))  +
        xlab("EGF duration (sec)") +
        # scale_x_discrete(
        #   "EGF treatment duration\n(sec)",
        #   drop = FALSE,
        #   labels = c("0", "10", "20", "30", "40", "50", "60", "70", "80")) +
          # limits = c("untreated", "1 min", "3 min", "5 min", "15 min"),
          # labels = c("0", "1", "3", "5", "15")) +
        # theme(axis.text.x = element_text(angle = -90, hjust = 0.5, vjust = 0.5)) +
        theme(
          strip.background = element_rect(fill = adjustcolor(strip_color, alpha.f = 1)),
          # strip.background = element_rect(fill = adjustcolor( strip_color, alpha.f = 0.3)),
          strip.text = element_text(size = 16, face = "bold", family = "sans", color = strip_text_color)) #+
        # expand_limits(y = c(15, 35)) 
      
      # save the plot as PNG & PDF
      ggsave(filename = file.path(output_dir, paste0(gene_var, "_EGF100nM.png")), plot = p, width = 8, height = 8, scale = 0.4)
      ggsave(filename = file.path(output_dir, paste0(gene_var, "_EGF100nM.pdf")), plot = p, width = 8, height = 8, scale = 0.4)
      
    } else { 
      message (paste("Gene", gene_var, "not found in the dataframe."))
      
      }
    }
}

# "#FFF5F0" "#FEE0D2" "#FCBBA1" "#FC9272" "#FB6A4A" "#EF3B2C" "#CB181D" "#A50F15" "#67000D"
#  "#FFF7FB" "#ECE7F2" "#D0D1E6" "#A6BDDB" "#74A9CF" "#3690C0" "#0570B0" "#045A8D" "#023858"
#   "#FCFBFD" "#EFEDF5" "#DADAEB" "#BCBDDC" "#9E9AC8" "#807DBA" "#6A51A3" "#54278F" "#3F007D"
#   brewer.pal(15, "Greys")
#   "#FFFFFF" "#F0F0F0" "#D9D9D9" "#BDBDBD" "#969696" "#737373" "#525252" "#252525" "#000000"


# cluster = c('1' = "#081D58", '2' = "#225EA8", '3'= "#41B6C4", '4' = "#C7E9B4"),
```


###__ii. run function to plot many PNGs
####___i.  cluster 1
```{r}

lineplot_gene_psite_function( vector_gene_refs = (EGF100nM_longer %>% distinct(gene_ref))$gene_ref,
                              strip_text_color = "white",
                             dataframe = EGF100nM_longer, output_dir = "output/MainFig3to6_SuppFig9to15/Reddy2016Comparison/", strip_color = "black")


```


### total pY sites
```{r}
net_pY_sites_Reddy2016_EGF100nM <- EGF100nM %>% 
  distinct(protein_short, sites)

write_csv(x = net_pY_sites_Reddy2016_EGF100nM, file = "modified_data/MainFig3to6_SuppFig9to15/Reddy2016Comparison/net_pY_sites_Reddy2016_EGF100nM.csv")
```

#### . . id mapping
```{r}
id_mapping <- read_csv(file = "raw_data/MainFig3to6_SuppFig9to15/ID_mapping_20250424.csv") %>% 
  group_by(gene) %>% 
  mutate(
    reference_number = row_number()) %>% 
  filter(reference_number == min(reference_number)) %>% 
  ungroup() %>% 
  rename(protein_short = from) %>% 
  select(-reference_number)

#append modern gene and reference names for matching to my database:

net_pY_sites_Reddy2016_EGF100nM_idmapped <- net_pY_sites_Reddy2016_EGF100nM %>% 
  left_join(id_mapping, by = c("protein_short")) %>% 
  mutate(psite = str_remove_all(sites, "[p]"))
```
##compare to our data
opened in the same document, otherwise will need to write and read in the file.
```{r}
our_data <- psite_diff_abundance_gene_2xDB_imputed_forvolcano %>% 
  filter(my_data == TRUE) %>% 
  distinct(gene, psite, gene_ref, reference, my_data)

comparison <- our_data %>% 
  full_join(y = net_pY_sites_Reddy2016_EGF100nM_idmapped %>% mutate(Reddy = TRUE), by = c("reference", "gene", "psite")) %>% 
  mutate(
    overlap  = case_when(
      my_data == TRUE & Reddy == TRUE ~ "both", 
      my_data == TRUE & is.na(Reddy) ~ "this\nstudy",
      is.na(my_data) & Reddy == TRUE ~ "Reddy et al.\n2016"))

comparison_summary <- comparison %>% group_by(overlap) %>% summarize(n_pYsites = n()) %>% ungroup()

#plot ----------------------
plot_comparison_Reddy_vs_Us <-ggplot(data = comparison %>% group_by(overlap) %>% summarize(n_pYsites = n()) %>% ungroup()) +
  geom_col(mapping = aes(x = overlap, y = n_pYsites), width = 0.8) +
  geom_text(mapping = aes(x = overlap, y = n_pYsites - 25, label = paste0(n_pYsites)), color = "white", fontface = "bold", size = 4) +
  alexis_theme() +
  theme(
    axis.text.x = element_text(size = 10, vjust = 1, hjust = 0.5) ) +
  ylab("unique pY sites") +
  xlab("intersection") #+
  # scale_y_continuous(limits = c(0, 700), breaks = c(seq(0,700, 100)), expand = c(0,0))

plot_comparison_Reddy_vs_Us

ggsave(filename = "output/MainFig3to6_SuppFig9to15/Reddy2016Comparison/plot_comparison_Reddy_vs_Us_EGF100nM.png", plot = plot_comparison_Reddy_vs_Us, width = 7, height = 8, scale = 0.4)
```


###___detected vs detected
```{r}
our_data_detected <- ascore_fasta_protein_mod_loc_allobservations %>%
  rename(psite = mod_protein_location) %>% 
  distinct(reference, psite) %>% 
  filter(grepl("Y", psite) == TRUE) %>% 
  mutate(my_data = TRUE) %>% 
  distinct(reference,psite, my_data)

comparison_detected <- our_data_detected %>% 
  full_join(y = net_pY_sites_Reddy2016_EGF100nM_idmapped %>% mutate(Reddy = TRUE), by = c("reference", "psite")) %>% 
  mutate(
    overlap  = case_when(
      my_data == TRUE & Reddy == TRUE ~ "both", 
      my_data == TRUE & is.na(Reddy) ~ "this\nstudy",
      is.na(my_data) & Reddy == TRUE ~ "Reddy et al.\n2016"))

comparison_summary_detected <- comparison_detected %>% group_by(overlap) %>% summarize(n_pYsites = n()) %>% ungroup()

#plot ----------------------
plot_comparison_Reddy_vs_Us_detected <-ggplot(data = comparison_detected %>% group_by(overlap) %>% summarize(n_pYsites = n()) %>% ungroup()) +
  geom_col(mapping = aes(x = overlap, y = n_pYsites), width = 0.8) +
  geom_text(mapping = aes(x = overlap, y = n_pYsites - 25, label = paste0(n_pYsites)), color = "white", fontface = "bold", size = 4) +
  alexis_theme() +
  theme(
    axis.text.x = element_text(size = 10, vjust = 1, hjust = 0.5) ) +
  ylab("unique pY sites") +
  xlab("intersection") #+
  # scale_y_continuous(limits = c(0, 700), breaks = c(seq(0,700, 100)), expand = c(0,0))

plot_comparison_Reddy_vs_Us_detected

ggsave(filename = "output/MainFig3to6_SuppFig9to15/Reddy2016Comparison/plot_comparison_Reddy_vs_Us_EGF100nM_detected.png", plot = plot_comparison_Reddy_vs_Us_detected, width = 7, height = 8, scale = 0.4)
```
We also compared our data to a time course study of early EGF signaling in MCF-10A cells that used a combination of three anti-pY antibodies (4G10, PY-100, PT-66) and TMT multiplexing for quantification12. Reddy et al. 2016 quantified 159 pY sites from 100 nM EGF stimulation of MCF-10A cells measured at 10s resolution between 0 and 80 seconds. Compared to our set of 1651 quantified pY sites, we observed decent agreement between studies, in which 91 pY sites were shared and 68 pY sites were unique to Reddy et al. 2016 while 1560 were unique to this study. 

Label-free analyses are cheaper but often struggle to compete with isobaric labeling on sensitivity  due the signal amplification and data completeness. Still, our method quantified 10-fold pY sites without expensive TMT reagent.  


###_direction of change

